Model Finding for (Co)recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-05 tex/main.tex
tuned algo

\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv_0( t\typ{\tau},\, p )$ : 
 \begin{itemize}
   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$:
    \begin{itemize}
      \item[] $\mathsf{let}$ $( r_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
      \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
      \item[] $\mathsf{let}$ $\chi = \bigwedge \Chi$;
      \item[] $\mathsf{if}$ $p = \ptrue$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n,\, \Chi )$
      \end{itemize}
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( r_1,\, D_1 ) = \conv_0( t_1,\, p )$;
      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
      \begin{itemize}
        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
      \end{itemize}
      \item[] else
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1,\, \forall \vec x.\; D_1 )$
      \end{itemize}
    \end{itemize}
   \item[] $\mathsf{else}$
   \begin{itemize}
     \item[] $\mathsf{return}$ $( t,\, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ),\, f \in \sfundefs{\Sigma} \} )$
   \end{itemize}
 \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv( \varphi )$ : 
 \begin{itemize}
   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$;
   %\item[] $\mathsf{assert}( D = \emptyset )$;
   \item[] $\mathsf{return}$ $\psi$
  \end{itemize}
\end{enumerate}

@@

If $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it
recursively calls $\conv$ on its body with the same polarity. If $t$ is a
function definition of a function in $\sfundefs{\Sigma}$, we instead construct
a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{\con{f}}( y )$.
In this case, since function definitions are a top-level conjunct, by case
analysis on the return values of $\conv$, we know that $D_1$ is empty. In the
case that $t$ is not a function definition, the resulting quantified formula is
reconstructed with the term $r_1$ returned by the recursive call, and a
quantified prefix is appended to the formula $D_1$ returned by the recursive
call.

>>>>

\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv_0( t\typ{\tau},\, p )$ : 
 \begin{itemize}
   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$:
    \begin{itemize}
      \item[] $\mathsf{let}$ $( t'_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
      \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
      \item[] $\mathsf{let}$ $\chi = \bigwedge \Chi$;
      \item[] $\mathsf{if}$ $p = \ptrue$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \emptyset )$
      \end{itemize}
      \item[] $\mathsf{else}$
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n),\, \Chi )$
      \end{itemize}
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( t'_1,\, D_1 ) = \conv_0( t_1,\, p )$;
      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
      \begin{itemize}
        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( t'_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
      \end{itemize}
      \item[] else
      \begin{itemize}
        \item[] $\mathsf{return}$ $( \forall \vec x.\; t'_1,\, \forall \vec x.\; D_1 )$
      \end{itemize}
    \end{itemize}
   \item[] $\mathsf{else}$
   \begin{itemize}
     \item[] $\mathsf{return}$ $( t,\, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ),\, f \in \sfundefs{\Sigma} \} )$
   \end{itemize}
 \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\item[\ ] 
$\conv( \varphi )$ : 
 \begin{itemize}
   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$;
   %\item[] $\mathsf{assert}( D = \emptyset )$;
   \item[] $\mathsf{return}$ $\psi$
  \end{itemize}
\end{enumerate}

@@

If $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it
recursively calls $\conv$ on its body with the same polarity. If $t$ is a
function definition of a function in $\sfundefs{\Sigma}$, we instead construct
a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
and replace all occurrences of $\vec x$ in $t'_1$ with $\vecfarg{\con{f}}( y )$.
In this case, since function definitions are a top-level conjunct, by case
analysis on the return values of $\conv$, we know that $D_1$ is empty. In the
case that $t$ is not a function definition, the resulting quantified formula is
reconstructed with the term $t'_1$ returned by the recursive call, and a
quantified prefix is appended to the formula $D_1$ returned by the recursive
call.
